Rates of true polar wander in convecting planets

### 2015-09-08 tpw_rate.tex
consistently use mathbf

\begin{equation}
\frac{d {\bf H}}{dt} + {\bf \Omega} \times {\bf H} = 0
\label{eq:euler}
\end{equation}
known as Euler's equations, where ${\bf H} = {\bf I} \cdot {\bf \Omega}$ is the angular momentum vector, $\bf I$ is the moment of inertia tensor, and $\bf \Omega$ is the angular velocity.
On a dynamic planet $\bf I$ may be a function of time, so to conserve angular momentum $\bf \Omega$ must be also vary with time.
In this case Equation~\eqref{eq:euler} is known as the Liouville equation \citep[e.g.][]{munk1960rotation}.
For a slowly convecting fluid such as a planetary mantle the inertial term $\partial {\bf H} / \partial t$ is negligible, so we may solve the simplified quasistatic equations
\begin{equation}
{\bf \Omega } (t)\times ( {\bf I}(t) \cdot {\bf \Omega }(t) ) = 0
\label{eq:quasistatic_liouville}
\end{equation}
Equation~\eqref{eq:quasistatic_liouville} indicates that ${\bf \Omega}$ and ${\bf H}$ are parallel, so a solution for ${\bf \Omega}(t)$ is equivalent to solving an eigenvalue problem for ${\bf I}(t)$, where the eigenvectors correspond to the principal axes and the eigenvalues correspond to the principal moments (where the most stable orientation of the planet corresponds to rotating about the largest principal axis).
In practice this eigenvalue approach has been often used in previous studies for computing the spin axis of a planet \citep[e.g.][]{steinberger1997changes, roberts2007cause}.
The moment of inertia tensor in~\eqref{eq:quasistatic_liouville} includes all contributions to the mass structure
of the planet, including the spherically symmetric mass distribution, rotational deformation, deformation due to self gravity, internal and surface density anomalies, and surface deflections due to density anomalies.
Here we are interested in contributions from mantle convection, so we neglect other processes.

For mantle convection problems the moment of inertia tensor is commonly separated into three parts \citep{ricard1993polar}:
\begin{equation}
I_{ij}(t) = I_0 \delta_{ij} + J_{ij}(t) + E_{ij}(t)
\label{eq:separation}
\end{equation}
where $I_0$ is the spherically symmetric reference moment, $J_{ij}$ is the contribution due to rotational deformation, and $E_{ij}$ is the contribution due to internal density anomalies, as well as the surface deflections caused by them.
If we plug this decomposition into Equation~\eqref{eq:quasistatic_liouville} the spherically symmetric part $I_0 \delta_{ij}$ drops out, and we are left with
\begin{equation}
{\bf \Omega} \times ({\bf J} \cdot {\bf \Omega}) = -{\bf \Omega } \times ( {\bf E} \cdot {\bf \Omega} )
\label{eq:disequilibrium}
\end{equation}
This form of the quasistatic Liouville equation makes clear that the polar wander problem represents a balance between the mismatches of the convective part of the moment of inertia ($\bf E$) and the rotational deformation part of the moment of inertia ($\bf J$).
Our goal is to characterize this balance from a perspective of scaling and fluid dynamics.

@@

The term on the right-hand side of equation~\eqref{eq:disequilibrium} represents the moment of inertia due to internal density anomalies as well as the surface deflections due to them.
This, as well, may be parameterized using a viscoelastic Love number approach:
\begin{equation} 
{\bf E} = \left[ \delta(t) + k^L(t) \right] * {\bf C}
\end{equation}
where $k^L$ is an internal loading Love number representing the surface deflection due to density anomalies, and $C_{ij}$ is the moment of inertia without that response.
Frequently the simplification is made that the timescale of the surface response is quick compared to the true polar wander timescale, and we may use the fluid limit geoid kernels \citep[e.g.][]{richards1984geoid}:  
\begin{equation}
E_{ij} = (1+k^L_f) C_{ij}
\end{equation}
An alternative to the Love number formulism is to calculate to surface deflections directly using mantle convection simulations with a true free surface boundary condition.
A recent implementation of a free surface boundary condition in the CIG-sponsored mantle convection software \texttt{ASPECT} \citep{rose2015free}
permits more general treatments of mantle rheology.

@@

We are in a position to address the rates of true polar wander for a given convective moment $\bf E$.
A considerable simplification occurs if we neglect secular changes in the rotation rate, and just consider changes in direction of the pole ($d \Omega^2 / dt = 2 {\Omega_i} \dot{ \Omega}_i = 0$)
Substituting Equation~\eqref{eq:rotational_deformation} into Equation~\eqref{eq:disequilibrium} we find
\begin{equation}
\frac{k_f a^5 T_1}{3G}{\bf \Omega} \times {\bf \dot{\Omega} } = {\bf \Omega} \times \left( {\bf E \cdot \Omega} \right)
\label{eq:liouville}
\end{equation}
Introducing a unit vector $\mitbf{\omega} = {\bf \Omega}/\|{\bf \Omega} \|$ and using Equation~\eqref{eq:fluid_love} for $k_f$,  we may solve this equation for ${\bf \dot{\mitbf{\omega}} }$:
\begin{equation}
 \dot{\mitbf{\omega}}  = \frac{1}{(C-A)T_1} \left[ {\bf E \cdot \mitbf{\omega}} - \left( {\bf \mitbf{\omega} \cdot E \cdot \mitbf{\omega} } \right) \mitbf{\omega} \right]
\label{eq:liouville_ode}
\end{equation}
Note that the quantity in brackets is similar in form to the shear stress on a plane in classical elastostatics.
If we enter the coordinate system of the convective moment of inertia $\bf E$ with principal moments $\lambda_1 \le \lambda_2 \le \lambda_3$ and define the orientation of $\mitbf{\omega}$ with colatitude $\theta$ and longitude $\phi$ we can write this in a more illuminating form (after some tedious algebra):
\ifdetail

@@

Without loss of generality we may choose the coordinate system such that $\phi=0$:
\begin{equation}
\dot{\theta} = - \frac{1}{(C-A)T_1} (\lambda_3-\lambda_1) \sin{2 \theta}
\label{eq:simple_milankovitch}
\end{equation}
These equations are a version of what has been called the ``Milankovitch theorem'' \citep{munk1960rotation}.  
Written this way, it is clear that the important quantities are $\theta$ and the differences between the eigenvalues of the convective moment $\bf E$, both of which depend on the structure and dynamics of mantle convection.  
They represent, respectively, the size of convective anomalies in the moment of inertia tensor and the angular mismatch between the rotation axis and the principle axis of the convective moment.
Both $\theta$ and $\lambda_i$ depend upon the dynamics of the convecting planetary mantle.
In order to derive estimates for them, we must consider those dynamics.

@@

\begin{equation}
\nabla \cdot {\bf u} = 0
\label{eq:conserve_mass}
\end{equation}

\begin{equation}
\begin{aligned}
- \nabla P + \eta \nabla^2 {\bf u} =  \rho {\bf g} -  \rho{\bf \Omega \times \Omega \times r}
\label{eq:navier_stokes}
\end{aligned}
\end{equation}

\begin{equation}
\frac{\partial T}{\partial t} + {\bf u} \cdot \nabla T = \kappa \nabla^2 T
\label{eq:energy}
\end{equation}
where $\bf u$ is the velocity, $P$ is the pressure, and $T$ is the temperature
(see Table~\ref{tab:parameters} for definitions of the other parameters).
In addition we use the simple equation of state
\begin{equation}
\rho = \rho_0 \left( 1 - \alpha (T-T_0) \right)
\label{eq:eos}
\end{equation}
The other parameters are defined in Table \ref{tab:parameters}.
Note that here we retain the centrifugal term, which is normally either neglected or absorbed into a modified pressure.
Dimensional analysis of this system (cf. \citet{barenblatt1996scaling}) requires four nondimensional numbers to characterize it
(a fifth one, defined by the ratio of the length of day to a diffusion timescale, does not appear in the governing equations).
Convenient choices for these numbers are listed in Table \ref{tab:nondim}, along with approximate Earth-like values for them.

@@

\begin{equation}
\begin{aligned}
- \nabla P_0 =  \rho_0 {\bf g} -  \rho_0 {\bf \Omega \times \Omega \times r}
\label{eq:hydrostatic}
\end{aligned}
\end{equation}
Nondimensionalizing with the parameters in Table \ref{tab:nondim} and removing the reference state we find the nondimensional momentum equation:
\begin{equation}
\begin{aligned}
 - \nabla P^* + \; \nabla^2{\bf v} - \mathrm{Ra} \; T \; {\bf g} + \mathrm{Ra \; Fr}\; T \;{\mitbf{\omega} \times \mitbf{\omega} \times {\bf r}} = 0
\end{aligned}
\end{equation}

@@

\begin{table}
\centering
\caption{Nondimensionalizations for variables, primed variables denote dimensional variables}
\label{tab:nondim_convert}
\begin{tabular}{@{}lll}
$x^\prime$ &=& $R \;\; x$ \\
$t^\prime$ &=& $R^2/\kappa \;\; t$ \\
$T^\prime$ &=& $\Delta T \;\; T$ \\
$\rho^\prime$ &=& $\rho_0 \;\; \rho$\\
$v^\prime$ &=& $\kappa/R \;\; v$ \\
$P^\prime$ &=& $\eta \kappa/R^2 \;\; P$ \\
$\Omega^\prime$ &=& $\Omega_0 \;\; \Omega$ \\
$g^\prime$ &=& $g_0 \;\; g$
\end{tabular}
\end{table}

If we plug these nondimensionalizations into the governing equations we find that the incompressibility constraint remains unchanged.  
The temperature equation and equation of state become
\begin{equation}
\frac{\partial T}{\partial t} + \mathbf{v} \cdot \nabla T = \nabla^2 T
\end{equation}

@@

\begin{equation}
\begin{aligned}
 - & \Omega_0^2  \alpha  \Delta T T R \; {\bf \Omega \times \Omega \times r} = - \eta \kappa / R^2 \; \nabla P^* \\ 
&+ \eta \kappa / R^2 \; \nabla \cdot \left( \eta \epsilon ({\bf v}) \right) - \alpha \Delta T g_0 \; {\bf g}
\end{aligned}
\end{equation}

\begin{equation}
\begin{aligned}
 - & \mathrm{Ra \; Fr}\; T \; {\bf \Omega \times \Omega \times r} = - \nabla P^* \\ 
&+ \; \nabla \cdot \left( \eta \epsilon ({\bf v}) \right) - \mathrm{Ra} \; T \; {\bf g}
\end{aligned}
\end{equation}

@@

We may explicitly draw a connection between the moment of inertia of a rotating object and it's degree-two density structure.  The moment of inertia tensor may be written in index notation
\begin{equation}
I_{ij} = \int_V \rho \left( r_q r_q \delta_{ij} - r_i r_j \right) dV
\label{eq:inertia}
\end{equation}
where $\bf r$ is the Eulerian coordinate, $\rho$ is the density, and $V$ is the volume of the material.  
It is useful to enter the principal axes of the moment of inertia:
\begin{equation}
{\bf I} = \bf{1} \begin{bmatrix}
\lambda_1 \\ \lambda_2 \\ \lambda_3
\end{bmatrix} = 
{\bf 1} \begin{bmatrix}
\int_V \rho (y^2+z^z) dV\\
\int_V \rho (x^2+z^2) dV\\
\int_V \rho (x^2+y^2) dV
\end{bmatrix}
\end{equation}
where $\bf 1$ is the identity matrix, and $\lambda_1$, $\lambda_2$, and $\lambda_3$ are the principal moments.
From equation~\eqref{eq:milankovitch} we see that the important quantities are the differences between the 
principal moments, $(\lambda_3-\lambda_1)$, $(\lambda_3-\lambda_2)$ and $(\lambda_2-\lambda_1)$.
These quantities may be rewritten in terms of degree-two real sperical harmonics \citep[e.g.][]{dahlen1999theoretical}.
The relevant (fully normalized) harmonics are, in Cartesian coordinates:
\begin{equation}
\begin{aligned}
Y_{20} &= \frac{1}{4} \sqrt{\frac{ 5}{\pi}} \frac{ 2 z^2 - x^2 - y^2}{r^2} \\ 
Y_{22} &= \frac{1}{4} \sqrt{\frac{15}{\pi}} \frac{ x^2 - y^2}{r^2}
\end{aligned}
\end{equation}
Solving for $( \lambda_i - \lambda_j )$ in terms of these harmonics, we find
\ifdetail

>>>>

\begin{equation}
\frac{d \mathbf{H}}{dt} + \mitbf{\Omega} \times \mathbf{H} = 0
\label{eq:euler}
\end{equation}
known as Euler's equations, where $\mathbf{H} = \mathbf{I} \cdot \mitbf{\Omega}$ is the angular momentum vector, $\mathbf{I}$ is the moment of inertia tensor, and $\mitbf{\Omega}$ is the angular velocity.
On a dynamic planet $\mathbf{I}$ may be a function of time, so to conserve angular momentum $\mitbf{\Omega}$ must be also vary with time.
In this case Equation~\eqref{eq:euler} is known as the Liouville equation \citep[e.g.][]{munk1960rotation}.
For a slowly convecting fluid such as a planetary mantle the inertial term $\partial \mathbf{H} / \partial t$ is negligible, so we may solve the simplified quasistatic equations
\begin{equation}
\mitbf{\Omega } (t)\times ( \mathbf{I}(t) \cdot \mitbf{\Omega }(t) ) = 0
\label{eq:quasistatic_liouville}
\end{equation}
Equation~\eqref{eq:quasistatic_liouville} indicates that $\mitbf{\Omega}$ and $\mathbf{H}$ are parallel, so a solution for $\mitbf{\Omega}(t)$ is equivalent to solving an eigenvalue problem for $\mathbf{I}(t)$, where the eigenvectors correspond to the principal axes and the eigenvalues correspond to the principal moments (where the most stable orientation of the planet corresponds to rotating about the largest principal axis).
In practice this eigenvalue approach has been often used in previous studies for computing the spin axis of a planet \citep[e.g.][]{steinberger1997changes, roberts2007cause}.
The moment of inertia tensor in~\eqref{eq:quasistatic_liouville} includes all contributions to the mass structure
of the planet, including the spherically symmetric mass distribution, rotational deformation, deformation due to self gravity, internal and surface density anomalies, and surface deflections due to density anomalies.
Here we are interested in contributions from mantle convection, so we neglect other processes.

For mantle convection problems the moment of inertia tensor is commonly separated into three parts \citep{ricard1993polar}:
\begin{equation}
I_{ij}(t) = I_0 \delta_{ij} + J_{ij}(t) + E_{ij}(t)
\label{eq:separation}
\end{equation}
where $I_0$ is the spherically symmetric reference moment, $J_{ij}$ is the contribution due to rotational deformation, and $E_{ij}$ is the contribution due to internal density anomalies, as well as the surface deflections caused by them.
If we plug this decomposition into Equation~\eqref{eq:quasistatic_liouville} the spherically symmetric part $I_0 \delta_{ij}$ drops out, and we are left with
\begin{equation}
\mathbf{\Omega} \times (\mathbf{J} \cdot \mitbf{\Omega}) = -\mitbf{\Omega} \times ( \mathbf{E} \cdot \mitbf{\Omega} )
\label{eq:disequilibrium}
\end{equation}
This form of the quasistatic Liouville equation makes clear that the polar wander problem represents a balance between the mismatches
of the convective part of the moment of inertia ($\mathbf{E}$) and the rotational deformation part of the moment of inertia ($\mathbf{J}$).
Our goal is to characterize this balance from a perspective of scaling and fluid dynamics.

@@

The term on the right-hand side of equation~\eqref{eq:disequilibrium} represents the moment of inertia due to internal density anomalies as well as the surface deflections due to them.
This, as well, may be parameterized using a viscoelastic Love number approach:
\begin{equation} 
\mathbf{E} = \left[ \delta(t) + k^L(t) \right] * \mathbf{C}
\end{equation}
where $k^L$ is an internal loading Love number representing the surface deflection due to density anomalies, and $C_{ij}$ is the moment of inertia without that response.
Frequently the simplification is made that the timescale of the surface response is quick compared to the true polar wander timescale, and we may use the fluid limit geoid kernels \citep[e.g.][]{richards1984geoid}:  
\begin{equation}
E_{ij} = (1+k^L_f) C_{ij}
\end{equation}
An alternative to the Love number formulism is to calculate to surface deflections directly using mantle convection simulations with a true free surface boundary condition.
A recent implementation of a free surface boundary condition in the CIG-sponsored mantle convection software \texttt{ASPECT} \citep{rose2015free}
permits more general treatments of mantle rheology.

@@

We are in a position to address the rates of true polar wander for a given convective moment $\mathbf{E}$.
A considerable simplification occurs if we neglect secular changes in the rotation rate, and just consider changes in direction of the pole ($d \Omega^2 / dt = 2 {\Omega_i} \dot{ \Omega}_i = 0$)
Substituting Equation~\eqref{eq:rotational_deformation} into Equation~\eqref{eq:disequilibrium} we find
\begin{equation}
\frac{k_f a^5 T_1}{3G}\mitbf{\Omega} \times \dot{\mitbf{\Omega} } = \mitbf{\Omega} \times \left( \mathbf{E} \cdot \mitbf{\Omega} \right)
\label{eq:liouville}
\end{equation}
Introducing a unit vector $\mitbf{\omega} = \mitbf{\Omega}/\|\mitbf{\Omega} \|$ and using Equation~\eqref{eq:fluid_love} for $k_f$,  we may solve this equation for $\dot{\mitbf{\omega}}$:
\begin{equation}
 \dot{\mitbf{\omega}}  = \frac{1}{(C-A)T_1} \left[ \mathbf{E} \cdot \mitbf{\omega} - \left( \mitbf{\omega} \cdot \mathbf{E} \cdot \mitbf{\omega}  \right) \mitbf{\omega} \right]
\label{eq:liouville_ode}
\end{equation}
Note that the quantity in brackets is similar in form to the shear stress on a plane in classical elastostatics.
If we enter the coordinate system of the convective moment of inertia $\mathbf{E}$ with principal moments $\lambda_1 \le \lambda_2 \le \lambda_3$ and define the orientation of $\mitbf{\omega}$ with colatitude $\theta$ and longitude $\phi$ we can write this in a more illuminating form (after some tedious algebra):
\ifdetail

@@

Without loss of generality we may choose the coordinate system such that $\phi=0$:
\begin{equation}
\dot{\theta} = - \frac{1}{(C-A)T_1} (\lambda_3-\lambda_1) \sin{2 \theta}
\label{eq:simple_milankovitch}
\end{equation}
These equations are a version of what has been called the ``Milankovitch theorem'' \citep{munk1960rotation}.  
Written this way, it is clear that the important quantities are $\theta$ and the differences between the eigenvalues of the convective moment $\mathbf{E}$, both of which depend on the structure and dynamics of mantle convection.  
They represent, respectively, the size of convective anomalies in the moment of inertia tensor and the angular mismatch between the rotation axis and the principle axis of the convective moment.
Both $\theta$ and $\lambda_i$ depend upon the dynamics of the convecting planetary mantle.
In order to derive estimates for them, we must consider those dynamics.

@@

\begin{equation}
\nabla \cdot \mathbf{u} = 0
\label{eq:conserve_mass}
\end{equation}

\begin{equation}
\begin{aligned}
- \nabla P + \eta \nabla^2 \mathbf{u} =  \rho \mathbf{g} -  \rho \mitbf{\Omega} \times \mitbf{\Omega} \times \mathbf{r}
\label{eq:navier_stokes}
\end{aligned}
\end{equation}

\begin{equation}
\frac{\partial T}{\partial t} + \mathbf{u} \cdot \nabla T = \kappa \nabla^2 T
\label{eq:energy}
\end{equation}
where $\mathbf{u}$ is the velocity, $P$ is the pressure, and $T$ is the temperature
(see Table~\ref{tab:parameters} for definitions of the other parameters).
In addition we use the simple equation of state
\begin{equation}
\rho = \rho_0 \left( 1 - \alpha (T-T_0) \right)
\label{eq:eos}
\end{equation}
The other parameters are defined in Table \ref{tab:parameters}.
Note that here we retain the centrifugal term, which is normally either neglected or absorbed into a modified pressure.
Dimensional analysis of this system (cf. \citet{barenblatt1996scaling}) requires four nondimensional numbers to characterize it
(a fifth one, defined by the ratio of the length of day to a diffusion timescale, does not appear in the governing equations).
Convenient choices for these numbers are listed in Table \ref{tab:nondim}, along with approximate Earth-like values for them.

@@

\begin{equation}
\begin{aligned}
- \nabla P_0 =  \rho_0 \mathbf{g} -  \rho_0 \mitbf{\Omega} \times \mitbf{\Omega} \times \mathbf{r}
\label{eq:hydrostatic}
\end{aligned}
\end{equation}
Nondimensionalizing with the parameters in Table \ref{tab:nondim} and removing the reference state we find the nondimensional momentum equation:
\begin{equation}
\begin{aligned}
 - \nabla P^* + \; \nabla^2 \mathbf{v} - \mathrm{Ra} \; T \; \mathbf{g} + \mathrm{Ra \; Fr}\; T \;{\mitbf{\omega} \times \mitbf{\omega} \times \mathbf{r}} = 0
\end{aligned}
\end{equation}

@@

\begin{table}
\centering
\caption{Nondimensionalizations for variables, primed variables denote dimensional variables}
\label{tab:nondim_convert}
\begin{tabular}{@{}lll}
$x^\prime$ &=& $R \;\; x$ \\
$t^\prime$ &=& $R^2/\kappa \;\; t$ \\
$T^\prime$ &=& $\Delta T \;\; T$ \\
$\rho^\prime$ &=& $\rho_0 \;\; \rho$\\
$u^\prime$ &=& $\kappa/R \;\; u$ \\
$P^\prime$ &=& $\eta \kappa/R^2 \;\; P$ \\
$\Omega^\prime$ &=& $\Omega_0 \;\; \Omega$ \\
$g^\prime$ &=& $g_0 \;\; g$
\end{tabular}
\end{table}

If we plug these nondimensionalizations into the governing equations we find that the incompressibility constraint remains unchanged.  
The temperature equation and equation of state become
\begin{equation}
\frac{\partial T}{\partial t} + \mathbf{u} \cdot \nabla T = \nabla^2 T
\end{equation}

@@

\begin{equation}
\begin{aligned}
 - & \Omega_0^2  \alpha  \Delta T T R \; \mitbf{\Omega} \times \mitbf{\Omega} \times \mathbf{r} = - \eta \kappa / R^2 \; \nabla P^* \\ 
&+ \eta \kappa / R^2 \; \nabla \cdot \left( \eta \epsilon (\mathbf{u}) \right) - \alpha \Delta T g_0 \; \mathbf{g}
\end{aligned}
\end{equation}

\begin{equation}
\begin{aligned}
 - & \mathrm{Ra \; Fr}\; T \; \mitbf{\Omega} \times \mitbf{\Omega} \times \mathbf{r} = - \nabla P^* \\ 
&+ \; \nabla \cdot \left( \eta \epsilon (\mathbf{u}) \right) - \mathrm{Ra} \; T \; \mathbf{g}
\end{aligned}
\end{equation}

@@

We may explicitly draw a connection between the moment of inertia of a rotating object and it's degree-two density structure.  The moment of inertia tensor may be written in index notation
\begin{equation}
I_{ij} = \int_V \rho \left( r_q r_q \delta_{ij} - r_i r_j \right) dV
\label{eq:inertia}
\end{equation}
where $\mathbf{r}$ is the Eulerian coordinate, $\rho$ is the density, and $V$ is the volume of the material.  
It is useful to enter the principal axes of the moment of inertia:
\begin{equation}
\mathbf{I} = \mathbf{1} \begin{bmatrix}
\lambda_1 \\ \lambda_2 \\ \lambda_3
\end{bmatrix} = 
\mathbf{1} \begin{bmatrix}
\int_V \rho (y^2+z^z) dV\\
\int_V \rho (x^2+z^2) dV\\
\int_V \rho (x^2+y^2) dV
\end{bmatrix}
\end{equation}
where $\mathbf{1}$ is the identity matrix, and $\lambda_1$, $\lambda_2$, and $\lambda_3$ are the principal moments.
From equation~\eqref{eq:milankovitch} we see that the important quantities are the differences between the 
principal moments, $(\lambda_3-\lambda_1)$, $(\lambda_3-\lambda_2)$ and $(\lambda_2-\lambda_1)$.
These quantities may be rewritten in terms of degree-two real sperical harmonics \citep[e.g.][]{dahlen1999theoretical}.
The relevant (fully normalized) harmonics are, in Cartesian coordinates:
\begin{equation}
\begin{aligned}
Y_{20} &= \frac{1}{4} \sqrt{\frac{ 5}{\pi}} \frac{ 2 z^2 - x^2 - y^2}{r^2} \\ 
Y_{22} &= \frac{1}{4} \sqrt{\frac{15}{\pi}} \frac{ x^2 - y^2}{r^2}
\end{aligned}
\end{equation}
Solving for $( \lambda_i - \lambda_j )$ in terms of these harmonics, we find
\ifdetail
